---
title: "Cold Springs Orchard 2021 Infection period analysis"
author: "Evan Krause"
output:
  pdf_document: default
  html_document: default
date: '2022-04-17'
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

```{r}
library("reshape2")
library("dplyr")
library("dtplyr")
library("data.table")
library("psych")
library("ggplot2")
```

```{r}
infect <- data.table(read.csv("Infection_event_2021.csv"))
sensors <- data.table(read.csv("sensors_wd.csv"))
names(infect)[1:6] = c(
  "date",
  "infect_events",
  "avg_wet_hours_temp",
  "wet_hours",
  "humid_hours(90+)",
  "rain_amt_in"
)
```

```{r}
infect_days <-
  infect %>% filter(infect$infect_events == "yes" |
                      infect$infect_events == "combined")
```

```{r}
## North sites
n_main <- sensors %>% filter(site == "n.main")
n_m <- sensors %>% filter(site == "n.m")
n_e <- sensors %>% filter(site == "n.e")
n_x <- sensors %>% filter(site == "n.x")

## South sites
s_main <- sensors %>% filter(site == "s.main")
s_a11 <- sensors %>% filter(site == "s.a11")
s_a16 <- sensors %>% filter(site == "s.a16")
s_g <- sensors %>% filter(site == "s.g")
s_y <- sensors %>% filter(site == "s.y")
```

# 2021 NEWA infection periods graphical exploration

## March 28th - 29th

```{r}
ggplot(sensors[date >= "2021-03-28" & date <= "2021-03-29", ],
       aes(date, temp)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site temperatures by day March 28th - 29th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site, margins = TRUE)

```

```{r}
ggplot(sensors[date >= "2021-03-28" & date <= "2021-03-29", ],
       aes(date, rh)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Relative humidity by day March 28th - 29th, 2021",
       y = "Relative humidity (%)") + 
  facet_grid(~ site)

```

```{r}
ggplot(sensors[date >= "2021-03-28" & date <= "2021-03-29", ],
       aes(date, wet)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Leaf wetness by day April 28th - 30th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

\newpage

## April 15th - 17th

```{r}
ggplot(sensors[date >= "2021-04-15" & date <= "2021-04-17", ],
       aes(date, temp)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site temperatures by day April 15th - 17th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)

```

```{r}
ggplot(sensors[date >= "2021-04-15" & date <= "2021-04-17", ],
       aes(date, rh)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Relative humidity by day April 15th - 17th, 2021",
       y = "Relative humidity (%)") + 
  facet_grid(~ site)

```

```{r}
ggplot(sensors[date >= "2021-04-15" & date <= "2021-04-17", ],
       aes(date, wet)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Leaf Wetness by day April 28th - 30th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

\newpage

## April 28th - 30th

```{r}
ggplot(sensors[date >= "2021-04-28" & date <= "2021-04-30", ],
       aes(date, temp)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site temperatures by day April 28th - 30th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

```{r}
ggplot(sensors[date >= "2021-04-28" & date <= "2021-04-30", ],
       aes(date, rh)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Relative humidity by day April 28th - 30th, 2021",
       y = "Relative humidity (%)") + 
  facet_grid(~ site)

```

```{r}
ggplot(sensors[date >= "2021-04-28" & date <= "2021-04-30", ],
       aes(date, wet)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Leaf Wetness by day April 28th - 30th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

\newpage

## May 2nd - 5th

```{r}
ggplot(sensors[date >= "2021-05-02" & date <= "2021-05-05", ],
       aes(date, temp)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site temperatures by day May 2nd - 5th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

```{r}
ggplot(sensors[date >= "2021-05-02" & date <= "2021-05-05", ],
       aes(date, rh)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Relative humidity by day May 2nd - 5th, 2021",
       y = "Relative humidity (%)") + 
  facet_grid(~ site)

```

```{r}
ggplot(sensors[date >= "2021-05-02" & date <= "2021-05-02", ],
       aes(date, wet)) +
  geom_boxplot(alpha = 0.5) + 
  labs(title = "CSO site Leaf Wetness by day May 2nd - 5th, 2021",
       y = "Temperature degF") + 
  facet_grid(~ site)
```

\newpage

# ANOVA and Post-Hoc
## By Infection period

### March 28th - 29th

#### Temperature
```{r}
temp_fit_03_28 <- aov(temp~site, data = subset(sensors, date >= "2021-03-28" & date <= "2021-03-29"))
anova(temp_fit_03_28)
temp_resid_03_28 <- resid(temp_fit_03_28)
hist(temp_resid_03_28,
     main = "March 28th - 29th Temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_03_28)
TukeyHSD(temp_fit_03_28, ordered = TRUE)
plot(TukeyHSD(temp_fit_03_28, ordered = TRUE))
```

\newpage

#### Relative Humidity

```{r}
rh_fit_03_28 <- aov(rh~site, data = subset(sensors, date >= "2021-03-28" & date <= "2021-03-29"))
anova(rh_fit_03_28)
rh_fit_resid_03_28 <- resid(rh_fit_03_28)
hist(rh_fit_resid_03_28,
     main = "March 28th - 29th Relative humidity anova",
     xlab = "Residuals")
qqnorm(rh_fit_resid_03_28)
TukeyHSD(rh_fit_03_28, ordered = TRUE)
plot(TukeyHSD(rh_fit_03_28, ordered = TRUE))
```

\newpage

#### Leaf Wetness

```{r}
wet_fit_03_28 <- aov(wet~site, data = subset(sensors, date >= "2021-03-28" & date <= "2021-03-29"))
anova(wet_fit_03_28)
wet_fit_resid_03_28 <- resid(wet_fit_03_28)
hist(wet_fit_resid_03_28,
     main = "March 28th - 29th Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_fit_resid_03_28)
TukeyHSD(wet_fit_03_28, ordered = TRUE)
plot(TukeyHSD(wet_fit_03_28, ordered = TRUE))
```

\newpage

### April 15th - 17th

#### Temperature
```{r}
temp_fit_04_15 <- aov(temp~site, data = subset(sensors, date >= "2021-04-15" & date <= "2021-04-17"))
anova(temp_fit_04_15)
temp_resid_04_15 <- resid(temp_fit_04_15)
hist(temp_resid_04_15,
     main = "April 15th - 17th Temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_04_15)
TukeyHSD(temp_fit_04_15, ordered = TRUE)
plot(TukeyHSD(temp_fit_04_15, ordered = TRUE))
```

\newpage

#### Relative Humidity

```{r}
rh_fit_04_15 <- aov(rh~site, data = subset(sensors, date >= "2021-04-15" & date <= "2021-04-17"))
anova(rh_fit_04_15)
rh_resid_04_15 <- resid(rh_fit_04_15)
hist(rh_resid_04_15,
     main = "April 15th - 17th Relative Humidity anova",
     xlab = "Residuals")
qqnorm(rh_resid_04_15)
TukeyHSD(rh_fit_04_15, ordered = TRUE)
plot(TukeyHSD(rh_fit_04_15, ordered = TRUE))
```

\newpage

#### Leaf Wetness

```{r}
wet_fit_04_15 <- aov(wet~site, data = subset(sensors, date >= "2021-04-15" & date <= "2021-04-17"))
anova(wet_fit_04_15)
wet_resid_04_15 <- resid(wet_fit_04_15)
hist(wet_resid_04_15,
     main = "April 15th - 17th Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_resid_04_15)
TukeyHSD(wet_fit_04_15, ordered = TRUE)
plot(TukeyHSD(wet_fit_04_15, ordered = TRUE))
```

\newpage

### April 28th - 30th

#### Temperature
```{r}
temp_fit_04_28 <- aov(temp~site, data = subset(sensors, date >= "2021-04-28" & date <= "2021-04-30"))
anova(temp_fit_04_28)
temp_resid_04_28 <- resid(temp_fit_04_28)
hist(temp_resid_04_28,
     main = "April 28th - 30th Temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_04_28)
TukeyHSD(temp_fit_04_28, ordered = TRUE)
plot(TukeyHSD(temp_fit_04_28, ordered = TRUE))
```

\newpage

#### Relative Humidity

```{r}
rh_fit_04_28 <- aov(rh~site, data = subset(sensors, date >= "2021-04-28" & date <= "2021-04-30"))
anova(rh_fit_04_28)
rh_resid_04_28 <- resid(rh_fit_04_28)
hist(rh_resid_04_28,
     main = "April 28th - 30th Relative Humidity anova",
     xlab = "Residuals")
qqnorm(rh_resid_04_28)
TukeyHSD(rh_fit_04_28, ordered = TRUE)
plot(TukeyHSD(rh_fit_04_28, ordered = TRUE))
```

\newpage

#### Leaf Wetness

```{r}
wet_fit_04_28 <- aov(wet~site, data = subset(sensors, date >= "2021-04-28" & date <= "2021-04-30"))
anova(wet_fit_04_28)
wet_resid_04_28 <- resid(wet_fit_04_15)
hist(wet_resid_04_28,
     main = "April 28th - 30th Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_resid_04_28)
TukeyHSD(wet_fit_04_28, ordered = TRUE)
plot(TukeyHSD(wet_fit_04_28, ordered = TRUE))
```

\newpage

### March 2nd - 5th


#### Temperature
```{r}
temp_fit_05_02 <- aov(temp~site, data = subset(sensors, date >= "2021-05-02" & date <= "2021-05-05"))
anova(temp_fit_05_02)
temp_resid_05_02 <- resid(temp_fit_05_02)
hist(temp_resid_05_02,
     main = "May 2nd - 5th Temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_05_02)
TukeyHSD(temp_fit_05_02, ordered = TRUE)
plot(TukeyHSD(temp_fit_05_02, ordered = TRUE))
```

\newpage

#### Relative Humidity

```{r}
rh_fit_05_02 <- aov(rh~site, data = subset(sensors, date >= "2021-05-02" & date <= "2021-05-05"))
anova(rh_fit_05_02)
rh_resid_05_02 <- resid(rh_fit_05_02)
hist(rh_resid_05_02,
     main = "May 2nd - 5th Relative Humidity anova",
     xlab = "Residuals")
qqnorm(rh_resid_05_02)
TukeyHSD(rh_fit_05_02, ordered = TRUE)
plot(TukeyHSD(rh_fit_05_02, ordered = TRUE))
```

\newpage

#### Leaf Wetness

```{r}
wet_fit_05_02 <- aov(wet~site, data = subset(sensors, date >= "2021-05-02" & date <= "2021-05-05"))
anova(wet_fit_05_02)
wet_resid_05_02 <- resid(wet_fit_05_02)
hist(wet_resid_05_02,
     main = "May 2nd - 5th Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_resid_05_02)
TukeyHSD(wet_fit_05_02, ordered = TRUE)
plot(TukeyHSD(wet_fit_05_02, ordered = TRUE))
```

\newpage

## By Month
### March 2021 Temperature by site

```{r}
temp_fit_03 <- aov(temp~site, data = subset(sensors, date >= "2021-03-01" & date <= "2021-03-31"))
anova(temp_fit_03)
temp_resid_03 <- resid(temp_fit_03)
hist(temp_resid_03,
     main = "April 2021 temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_03)
TukeyHSD(temp_fit_03, ordered = TRUE)
plot(TukeyHSD(temp_fit_03, ordered = TRUE))
```

\newpage

### March 2021 Relative Humidity by Site

```{r}
rh_fit_03 <- aov(rh~site, data = subset(sensors, date >= "2021-03-01" & date <= "2021-03-31"))
anova(rh_fit_03)
rh_fit_resid_03 <- resid(rh_fit_03)
hist(rh_fit_resid_03,
     main = "April 2021 Relative humidity anova",
     xlab = "Residuals")
qqnorm(rh_fit_resid_03)
TukeyHSD(rh_fit_03, ordered = TRUE)
plot(TukeyHSD(rh_fit_03, ordered = TRUE))
```

\newpage

### March 2021 Leaf Wetness by Site

```{r}
wet_fit_03 <- aov(wet~site, data = subset(sensors, date >= "2021-03-01" & date <= "2021-03-31"))
anova(wet_fit_03)
wet_fit_resid_03 <- resid(wet_fit_03)
hist(wet_fit_resid_03,
     main = "April 2021 Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_fit_resid_03)
TukeyHSD(wet_fit_03, ordered = TRUE)
plot(TukeyHSD(wet_fit_03, ordered = TRUE))
```

\newpage

### April 2021 Temperature by site

```{r}
temp_fit_04 <- aov(temp~site, data = subset(sensors, date >= "2021-04-01" & date <= "2021-04-30"))
anova(temp_fit_04)
temp_resid_04 <- resid(temp_fit_04)
hist(temp_resid_04,
     main = "April 2021 temperature anova",
     xlab = "Residuals")
qqnorm(temp_resid_04)
TukeyHSD(temp_fit_04, ordered = TRUE)
plot(TukeyHSD(temp_fit_04, ordered = TRUE))
```

\newpage

### April 2021 Relative Humidity by Site

```{r}
rh_fit_04 <- aov(rh~site, data = subset(sensors, date >= "2021-05-01" & date <= "2021-05-30"))
anova(rh_fit_04)
rh_fit_resid_04 <- resid(rh_fit_04)
hist(rh_fit_resid_04,
     main = "April 2021 Relative humidity anova",
     xlab = "Residuals")
qqnorm(rh_fit_resid_04)
TukeyHSD(rh_fit_04, ordered = TRUE)
plot(TukeyHSD(rh_fit_04, ordered = TRUE))
```

\newpage

### April 2021 Leaf Wetness by Site

```{r}
wet_fit_04 <- aov(wet~site, data = subset(sensors, date >= "2021-04-01" & date <= "2021-04-30"))
anova(wet_fit_04)
wet_fit_resid_04 <- resid(wet_fit_04)
hist(wet_fit_resid_04,
     main = "April 2021 Leaf Wetness anova",
     xlab = "Residuals")
qqnorm(wet_fit_resid_04)
TukeyHSD(wet_fit_04, ordered = TRUE)
plot(TukeyHSD(wet_fit_04, ordered = TRUE))
```

\newpage


### May 2021 Temperature by site

```{r}
temp_fit <- aov(temp~site, data = subset(sensors, date >= "2021-05-01" & date <= "2021-05-30"))
anova(temp_fit)
temp_resid <- resid(temp_fit)
hist(temp_resid,
     main = "May 2021 temperature anova",
     xlab = "Residuals")
TukeyHSD(temp_fit, ordered = TRUE)
plot(TukeyHSD(temp_fit, ordered = TRUE))
```

\newpage

### May 2021 Relative Humidity by Site

```{r}
rh_fit_05 <- aov(rh~site, data = subset(sensors, date >= "2021-05-01" & date <= "2021-05-30"))
anova(rh_fit_05)
rh_fit_05_resid <- resid(rh_fit_05)
hist(rh_fit_05_resid,
     main = "May 2021 Relative humidity anova",
     xlab = "Residuals")
TukeyHSD(rh_fit_05, ordered = TRUE)
plot(TukeyHSD(rh_fit_05, ordered = TRUE))
```

\newpage

### May 2021 Leaf Wetness by Site

```{r}
wet_fit_05 <- aov(wet~site, data = subset(sensors, date >= "2021-05-01" & date <= "2021-05-30"))
anova(wet_fit_05)
wet_fit_05_resid <- resid(wet_fit_05)
hist(wet_fit_05_resid,
     main = "May 2021 Relative humidity anova",
     xlab = "Residuals")
TukeyHSD(wet_fit_05, ordered = TRUE)
plot(TukeyHSD(wet_fit_05, ordered = TRUE))
```
