---
title: "2020-21 OrchardWatch sensor analysis"
author: "Evan Krause"
output: pdf_document
code_folding: "hide"
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r}
# req lib
library("reshape2")
library("dplyr")
library("ggplot2")
library("tidyr")
library("psych")
library("data.table")

# import sensor file
sensors <- data.table(read.csv('ow_ns.csv'))
names(sensors)[1] <- "datetime"

#Convert to long form
sensor_melt <- melt(sensors, na.rm = FALSE)
test_var = rep("rh_n-main", 5)
test_out = strsplit(test_var, split = "_") ## split para_reg_site form
unlist(test_out) ## simplify list structure to vector
matrix(unlist(test_out), ncol = 3, byrow = T) ## convert list to matrix

test_var <-
  as.character(sensor_melt$variable) ## define variable col as character
test_out = strsplit(test_var, split = "_") ## split para_reg_site form

para_frame <-
  data.table(matrix(unlist(test_out), ncol = 2, byrow = T)) ## convert to data frame
#para_frame <- para_frame %>% separate(X1, c("temp", "rh", "wet"))
names(para_frame) = c("para", "site")

sensor2 <-
  cbind(sensor_melt, para_frame) ## Combine melted data and site/para cols

#Convert datetime
sensor2$date <-
  format(as.POSIXct(sensor2$datetime, format = "%m/%d/%Y %H:%M"),
         "%Y/%m/%d")
sensor2$time <-
  format(as.POSIXct(sensor2$datetime, format = "%m/%d/%Y %H:%M"),
         "%H:%M")

#Convert date from char to date class and rename
date_frame <- data.table(as.Date.character(sensor2$date))
sensorFinal <- cbind(sensor2, date_frame)
names(sensorFinal)[8] <- "Date"
sensorFinal$date <- NULL

##Values separated by parameter and year
rh_vals <-
  subset(sensorFinal,
         sensorFinal$para == "rh")
rh_vals20 <- subset(rh_vals, rh_vals$Date <= "2021-03-26")
rh_vals21 <- subset(rh_vals, rh_vals$Date > "2021-03-26")

temp_vals <-
  subset(sensorFinal,
         sensorFinal$para == "temp")
temp_vals20 <- subset(temp_vals, temp_vals$Date <= "2021-03-26")
temp_vals21 <- subset(temp_vals, temp_vals$Date > "2021-03-26")

wet_vals <-
  subset(sensorFinal,
         sensorFinal$para >= "wet")
wet_vals20 <- subset(wet_vals, wet_vals$Date <= "2021-03-26")
wet_vals21 <- subset(wet_vals, wet_vals$Date > "2021-03-26")

## Define parameter frame
para_vals <-
  data.table(
    sensorFinal$Date,
    sensorFinal$time,
    sensorFinal$site,
    temp_vals$value,
    rh_vals$value,
    wet_vals$value
  )
names(para_vals) <- c("date", "time", "site", "temp", "rh", "wet")

para2 <- data.table(
  temp_vals$value,
    rh_vals$value,
    wet_vals$value
  )
```
```{r}
## North sites
n_main <- sensorFinal %>% filter(site == "n.main")
n_m <- sensorFinal %>% filter(site == "n.m")
n_e <- sensorFinal %>% filter(site == "n.e")
n_x <- sensorFinal %>% filter(site == "n.x")

## South sites
s_main <- sensorFinal %>% filter(site == "s.main")
s_a11 <- sensorFinal %>% filter(site == "s.a11")
s_a16 <- sensorFinal %>% filter(site == "s.a16")
s_g <- sensorFinal %>% filter(site == "s.g")
s_y <- sensorFinal %>% filter(site == "s.y")


# N-main
nmain_temp <- subset(n_main, para == "temp")
nmain_rh <- subset(n_main, para == "rh")
nmain_wet <- subset(n_main, para == "wet")

# N-M
nm_temp = subset(n_m, para == "temp")
nm_rh = subset(n_m, para == "rh")
nm_wet = subset(n_m, para == "wet")

# N-E
ne_temp = subset(n_e, para == "temp")
ne_rh <- subset(n_e, para == "rh")
ne_wet <- subset(n_e, para == "wet")

# N-X
nx_temp <- subset(n_x, para == "temp")
nx_rh <- subset(n_x, para == "rh")
nx_wet <- subset(n_x, para == "wet")

# S-main
smain_temp <- subset(s_main, para == "temp")
smain_rh <- subset(s_main, para == "rh")
smain_wet <- subset(s_main, para == "wet")

# S-A11
sa11_temp <- subset(s_a11, para == "temp")
sa11_rh <- subset(s_a11, para == "rh")
sa11_wet <- subset(s_a11, para == "wet")

# S-A16
sa16_temp <- subset(s_a16, para == "temp")
sa16_rh <- subset(s_a16, para == "rh")
sa16_wet <- subset(s_a16, para == "wet")

# S-Y
sy_temp <- subset(s_y, para == "temp")
sy_rh <- subset(s_y, para == "rh")
sy_wet <- subset(s_y, para == "wet")

# S-G
sg_temp <- subset(s_g, para == "temp")
sg_rh <- subset(s_g, para == "rh")
sg_wet <- subset(s_g, para == "wet")

```



# Numeric exploration


## values orchard-wide during first inoculum period

### Temperature
```{r}
"2020"
summary(temp_vals20$value)
"2021"
summary(temp_vals21$value)
```

### Relative humidity
```{r}
"2020"
summary(rh_vals20$value)
"2021"
summary(rh_vals21$value)
```

### Leaf wetness
```{r}
"2020"
summary(wet_vals20$value)
"2021"
summary(wet_vals21$value)
```
## Site Comparisons

### North - main vs South - main 2020
```{r}
"Temperature"
t.test(nmain_temp$value, smain_temp$value,
       data = subset(sensorFinal, sensorFinal$Date <= "2021-03-26"))
"Relative humidity"
t.test(nmain_rh$value, smain_rh$value,
       data = subset(sensorFinal, sensorFinal$Date <= "2021-03-26"))
"Leaf wetness"
t.test(nmain_wet$value, smain_wet$value,
       data = subset(sensorFinal, sensorFinal$Date <= "2021-03-26"))
```


### Relative humidity as a function of temperature and wetness (additive - North Main site)
```{r}
TukeyHSD(aov(nmain_rh$value ~ nmain_temp$value + nmain_wet$value))
anova(aov(nmain_temp$value ~ nmain_rh$value))

```


# Graphical exploration

## Temperature values by site
```{r}
par(mfrow = c(3,3))

hist(nmain_temp$value,
     main = "N-main",
     xlab = "Temperature degF")
hist(nx_temp$value,
     main = "N-X",
     xlab = "Temperature degF")
hist(ne_temp$value,
     main = "N-E",
     xlab = "Temperature degF")
hist(nm_temp$value,
     main = "N-M",
     xlab = "Temperature degF")
hist(smain_temp$value,
     main = "S-main",
     xlab = "Temperature degF")
hist(sa11_temp$value,
     main = "S-A11",
     xlab = "Temperature degF")
hist(sa16_temp$value,
     main = "S-A16",
     xlab = "Temperature degF")
hist(sy_temp$value,
     main = "S-Y",
     xlab = "Temperature degF")
hist(sg_temp$value,
     main = "S-G",
     xlab = "Temperature degF")

```


```{r}
boxplot(
  nmain_temp$value,
  nx_temp$value,
  ne_temp$value,
  nm_temp$value,
  smain_temp$value,
  sa11_temp$value,
  sa16_temp$value,
  sy_temp$value,
  sg_temp$value,
  ylab = "Temperature degF",
  main = "2020-2021 Temperature ranges",
  names = c(
    "N-ma",
    "N-X",
    "N-E",
    "N-M",
    "S-ma",
    "S-A11",
    "S-A16",
    "S-Y",
    "S-G"
  )
)
#label = c("N-main","N-X", "N-E", "N-M", "S-main", "S-A11", "S-A16", "S-Y", "S-G")
```



## Relative humidity values by site

```{r}
par(mfrow = c(3,3))

hist(nmain_rh$value,
     main = "N-main",
     xlab = "Temperature degF")
hist(nx_rh$value,
     main = "N-X",
     xlab = "Temperature degF")
hist(ne_rh$value,
     main = "N-E",
     xlab = "Temperature degF")
hist(nm_rh$value,
     main = "N-M",
     xlab = "Temperature degF")
hist(smain_rh$value,
     main = "S-main",
     xlab = "Temperature degF")
hist(sa11_rh$value,
     main = "S-A11",
     xlab = "Temperature degF")
hist(sa16_rh$value,
     main = "S-A16",
     xlab = "Temperature degF")
hist(sy_rh$value,
     main = "S-Y",
     xlab = "Temperature degF")
hist(sg_rh$value,
     main = "S-G",
     xlab = "Temperature degF")

```

```{r}
boxplot(
  nmain_rh$value,
  nx_rh$value,
  ne_rh$value,
  nm_rh$value,
  smain_rh$value,
  sa11_rh$value,
  sa16_rh$value,
  sy_rh$value,
  sg_rh$value,
  ylab = "Relative humidity (%)",
  main = "2020-2021 RH ranges",
  names = c(
    "N-ma",
    "N-X",
    "N-E",
    "N-M",
    "S-ma",
    "S-A11",
    "S-A16",
    "S-Y",
    "S-G"
  )
)
```


## Wetness values by site

```{r}

par(mfrow = c(3,3))

hist(nmain_wet$value,
     main = "N-main",
     xlab = "Wetness %")
hist(nx_wet$value,
     main = "N-X",
     xlab = "Wetness %")
hist(ne_wet$value,
     main = "N-E",
     xlab = "Wetness %")
hist(nm_wet$value,
     main = "N-M",
     xlab = "Wetness %")
hist(smain_wet$value,
     main = "S-main",
     xlab = "Wetness %")
hist(sa11_wet$value,
     main = "S-A11",
     xlab = "Wetness %")
hist(sa16_wet$value,
     main = "S-A16",
     xlab = "Wetness %")
hist(sy_wet$value,
     main = "S-Y",
     xlab = "Wetness %")
hist(sg_wet$value,
     main = "S-G",
     xlab = "Wetness %")

```

```{r}
boxplot(
  nmain_wet$value,
  nx_wet$value,
  ne_wet$value,
  nm_wet$value,
  smain_wet$value,
  sa11_wet$value,
  sa16_wet$value,
  sy_wet$value,
  sg_wet$value,
  ylab = "Leaf Wetness (%)",
  main = "2020-2021 Leaf Wetness ranges",
  names = c(
    "N-ma",
    "N-X",
    "N-E",
    "N-M",
    "S-ma",
    "S-A11",
    "S-A16",
    "S-Y",
    "S-G"
  )
)
```


## North sites RH ~ leaf wetness
```{r}
par(mfrow = c(2,2))

plot(nmain_wet$value ~ nmain_rh$value,
     subset = nmain_rh$value < 100)
plot(nx_wet$value ~ nx_rh$value,
     subset = nx_rh$value < 100)
plot(ne_wet$value ~ ne_rh$value,
     subset = ne_rh$value < 100)
plot(nm_wet$value ~ nm_rh$value,
     subset = nm_rh$value < 100)

```

